---
title: "Getting started with gwasrapidd"
author: ""
date: ""
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with gwasrapidd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

gwasrapidd is a REST API R client for the [GWAS
Catalog](https://www.ebi.ac.uk/gwas), allowing you to programatically download:

- Studies with `get_studies()`
- Associations with `get_associations()`
- Variants with `get_variants()`
- Traits with `get_traits()`

Each function returns an S4 object that encapsulates a set of tables (dataframes), that together, form a relational database for each object type.

## Installation

```{r, eval = FALSE}
# Install the development version from GitHub:
# install.packages("devtools")
devtools::install_github("ramiromagno/gwasrapidd")
```

## Cheatsheet

TODO

## Getting started

The GWAS Catalog is a service provided by the EMBL-EBI and NHGRI that offers a
manually curated and freely available database of published genome-wide
association studies (GWAS). The Catalog website and infrastructure is hosted by
the [EMBL-EBI](https://www.ebi.ac.uk).

There are three ways to access the Catalog database:

- Via the Web interface at
[https://www.ebi.ac.uk/gwas](https://www.ebi.ac.uk/gwas)
- By downloading a
database dump in TSV or RDF/OWL formats:
[https://www.ebi.ac.uk/gwas/docs/file-downloads](https://www.ebi.ac.uk/gwas/docs/file-downloads)
- Or through the GWAS Catalog REST API:
[https://www.ebi.ac.uk/gwas/rest/docs/api](https://www.ebi.ac.uk/gwas/rest/docs/api)

gwasrapidd facilitates the access to the Catalog via the Catalog REST API,
allowing you to programmatically retrieve data directly into R.



## Usage

The Catalog REST API provides four types of data objects for retrieval:

- Studies
- Associations
- Variants
- Traits

gwasrapidd provides four corresponding functions to get each of the object
types:

- `get_studies()`
- `get_associations()`
- `get_variants()`
- `get_traits()`


## Examples

### Searching for information about Diabetes mellitus

```{r}

```


